### 2024-05-02 doc/manuscript.Rmd
first commit

---
title: "Superior paper quality with R-Markdown: The OBVIOUS Study"
bibliography: refs.bib
csl: jama.csl
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
    keep_md: true
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      dpi = 600, 
                      fig.width = 11, 
                      fig.height = 6)

# load standard packages for RMD word docs
library(officer)
library(flextable)
library(gtsummary)
library(dplyr)
library(purrr)

# specify a rounding specification 
library(table.glue)

conflicted::conflict_prefer('as_flextable', winner = 'flextable')
conflicted::conflict_prefer('filter', winner = 'dplyr')

rspec <- round_spec() |>
  round_using_magnitude(digits = c(2, 1, 0),
                        breaks = c(1, 10, Inf))

# save it to options:
names(rspec) <- paste('table.glue', names(rspec), sep = '.')
options(rspec)

# default is double space 
# single spacing for title page, abstract, captions
single_par <- fp_par(line_spacing=1)

```

Your name here, your degree,^1^ 
Second in Command, MS,^1^ 
Minor Contributor, PhD,^2^ and 
The Mastermind, PhD^1^

^1^Your department, University, City, State. `r single_par`
^2^Department of Win, Overtly Pompous University, Irvine, CA. `r single_par`

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

**Correspondence to:** \
Your name, Degree \
Your Department \
Additional departmental details if needed \
Your university \
Address line 1 \
Address line 2 \
Address line 3 \
your phone number \
your email `r single_par`

**Word Count:** Abstract: ; Manuscript:

```{r data-prep, include=FALSE}

library(palmerpenguins)

data <- penguins

```

\newpage

# ABSTRACT

\newpage

# INTRODUCTION

Here's how to cite one thing.[@r_language]

Here's how to cite two or more things.[@ambrosius_design_2014; @sprint_research_group_randomized_2015]

# METHODS

# RESULTS

```{r results-prep}

n_penguins <- nrow(data)

mean_bill_length <- mean(data$bill_length_mm, na.rm = TRUE)

mean_mass <- mean(data$body_mass_g, na.rm = TRUE)
sd_mass <- sd(data$body_mass_g, na.rm = TRUE)

# make a short name for table.glue::table_value - this makes
# the inline reporting of results less complicated to read.
tv <- table.glue::table_value
tg <- table.glue::table_glue

```

Here's how to put some results into the text: A total of `r n_penguins` were included in the current analysis. The mean bill length was `r tv(mean_bill_length)`. The mean (standard deviation) body mass was `r tg("{mean_mass} ({sd_mass})")`

# DISCUSSION

\newpage

# REFERENCES

<div id="refs"></div>

\newpage

# ACKNOWLEDGMENTS

# FINANCIAL DISCLOSURE

```{r tbl_footers}

ckd_defn <- as_paragraph("Chronic Kidney Disease was defined by an estimated glomerular filtration rate <60 ml/min/1.73 m", as_sup(2L), " based on the 2021 CKD-EPI creatinine equation.")

```

```{r tbl_abbrvs}

abbrvs <- list(
  CI   = 'confidence interval',
  FI   = 'frailty index',
  SD   = 'standard deviation',
  IQR  = 'interquartile range',
  MoCA = 'Montreal cognitive assessment',
  EHR  = 'electronic health records',
  CVD  = 'cardiovascular disease'
)

abbrvs_paste <- function(strings){

  front <- 'Abbreviations'

  matter <- glue::glue_collapse(strings, sep = '; ', last = '; and ')

  paste(front, matter, sep = ': ')

}

abbrvs_write <- function(abbr){

  sorted <- abbr[sort(names(abbr))]

  strings <- map2_chr(names(sorted), sorted, paste, sep = ' = ')

  as_paragraph(abbrvs_paste(strings))

}

```

**Table 1: Penguin characteristics**

```{r}

tbl_summary(data, by = 'island') %>% 
 as_flex_table() %>% 
 theme_box() %>% 
 align(j = c(2,3,4), align = 'center')

```

\newpage

**Table 2: A more customized table**

```{r}

data %>% 
 group_by(species) %>% 
 slice(1:3) %>% 
 select(-body_mass_g, -sex, -year) %>% 
 mutate(across(where(is.numeric), table_value)) %>% 
 as_grouped_data(groups = 'species') %>% 
 as_flextable(hide_grouplabel = TRUE) %>% 
 set_header_labels(bill_length_mm = 'Bill length, mm',
                   bill_depth_mm = 'Bill depth, mm',
                   flipper_length_mm = 'Flipper length, mm') %>% 
 add_header_row(values = c(" ", " Penguin things"),
                colwidths = c(1, 3)) %>% 
 bg(i = ~!is.na(species), bg = 'grey90') %>% 
 width(width = 1.5) %>% 
 width(j = 1, width = 2) %>% 
 align(part = 'all', j = c(2,3,4), align = 'center') %>% 
 footnote(i = 2, j = 2, 
          part = 'header',
          value = as_paragraph("Bill length doesn't correlate with what you think it does."),
          ref_symbols = '1') %>% 
 footnote(i = 1, j = 1, 
          part = 'body',
          value = as_paragraph('Adelies are the best and they know it'), 
          ref_symbols = '2') %>% 
 add_footer_lines(values = abbrvs_write(abbrvs[c('CI', 'EHR', 'SD')])) %>% 
 add_footer_lines(values = "none of those abbreviations are in the table but they demonstrate how to include abbreviations using my helper functions")

```

<!-- new page not needed at the end of a block section -->

```{r}

margins_normal <- page_mar(bottom = 2.54 * 0.3937, 
                           top    = 2.54 * 0.3937, 
                           right  = 2.54 * 0.3937, 
                           left   = 2.54 * 0.3937,
                           gutter = 0)

block_section(
  prop_section(
    page_size = page_size(orient = "portrait"),
    type = "continuous",
    page_margins = margins_normal
  )
)

```

To make a landscape section, use the block_section code here. Note that the section will be applied to all content prior to the block reaching up to the most recent block.

```{r echo = TRUE}
block_section(
  prop_section(
    page_size = page_size(orient = "landscape"),
    page_margins = page_mar(bottom = 1/8, top = 1/8, 
                            right = 1/4, left = 1/4),
    type = "continuous"
  )
)
```

To switch back to regular portrait section, add another portrait block. Notice how the block below acts on all content above it up to the prior block.

```{r echo = TRUE}
block_section(
  prop_section(
    page_size = page_size(orient = "portrait"),
    type = "continuous",
    page_margins = margins_normal
  )
)
```

### 2024-05-02 R/dpp_exclude.R

@@ -1,0 +1,43 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param dpp_formatted
+dpp_exclude <- function(dpp_formatted) {
+
+ n_participants <- nrow(dpp_formatted)
+
+ e1 <- dpp_formatted %>%
+  filter(
+   # drop those who have diabetes already (according to current defn)
+   glucose_fasting_mgdl < 126,
+   hba1c_percent < 6.5 | is.na(hba1c_percent)
+  )
+
+ n_participants <- c(n_participants, nrow(e1))
+
+ e2 <- e1 %>%
+  filter(treatment != 'troglitazone') %>%
+  droplevels()
+
+ n_participants <- c(n_participants, nrow(e2))
+
+ exclusions <- enframe(n_participants) %>%
+  mutate(dropped = abs(c(0, diff(value))),
+         name = c("Study participants",
+                   "Pre-diabetic at baseline",
+                   "Randomized to placebo, metformin, or lifestyle"))
+
+
+ out <- list(included = e2,
+             excluded = setdiff(dpp_formatted, e2))
+
+ # assert no loss of participants
+ stopifnot(nrow(bind_rows(out)) == nrow(dpp_formatted))
+
+ attr(out, 'exclusions') <- exclusions
+
+ out
+
+}

### 2024-05-02 R/dpp_format.R

@@ -1,0 +1,482 @@
+
+dpp_format <- function(dpp_merged) {
+
+ dpp_merged %>%
+  transmute(
+
+   # Predictors ---------------------------------------------------------------
+
+   # Demographics ----
+
+   age_yrs = agerand,
+
+
+   treatment = factor(tolower(assign)),
+
+   sex = factor(sex,
+                levels = 1:2,
+                labels = c("male",
+                           "female")),
+
+   education = edugrp,
+
+   race_ethnicity = factor(race_eth,
+                           levels = 1:4,
+                           labels = c("caucasian",
+                                      "african_american",
+                                      "hispanic",
+                                      "other")),
+
+   # Body measurements ----
+
+   weight_lbs = weight,
+
+   bmi = bmi,
+
+   # Labs ----
+
+   insulin_fasting_uuml = i000,
+
+   insulin_30m_uuml = i030,
+
+   proinsulin_fasting_pm = pin,
+
+   glucose_fasting_mgdl = g000,
+
+   glucose_30m_mgdl = g030,
+
+   glucose_2hr_mgdl = g120,
+
+   chol_trig_mgdl = trig,
+
+   chol_total_mgdl = chol,
+
+   chol_hdl_mgdl = chdl,
+
+   chol_ldlv_mgdl = vldl,
+
+   chol_ldlc_mgdl = cldl,
+
+   chol_ldlb_mgdl = ldlb,
+
+   chol_ldlc_mgdl = ldlc,
+
+   chol_ldl_particle_size = ldlz,
+
+   chol_hdl_to_trig_ratio = chol_hdl_mgdl / chol_trig_mgdl,
+
+   creatinine_urine_mgdl = ucre,
+
+   hba1c_percent = hba1,
+
+   fibrinogen_mgdl = fibr,
+
+   crp_mgdl = crp,
+
+   albumin_urine_mgdl = ualb,
+
+   serum_ast_ul = sgot,
+
+   serum_alt_ul = sgpt,
+
+   serum_creatinine = crea,
+
+   serum_bicarb = hco3,
+
[…]

### 2024-05-02 R/dpp_impute.R

@@ -1,0 +1,36 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param dpp_data
+dpp_impute_miceranger <- function(dpp_data,
+                                  n_impute = 3,
+                                  n_iter = 5,
+                                  n_neighbors = 10,
+                                  n_trees = 250) {
+
+
+ preds <- setdiff(
+  x = names(dpp_data),
+  # don't use the outcome in imputation models b/c
+  # we can safely assume that the outcome won't be
+  # known when we impute missing values in testing data
+  y = c("time", "status")
+ )
+
+ vars <- vector(mode = 'list', length = length(preds))
+ names(vars) <- preds
+
+ for(v in seq_along(vars)) vars[[v]] <- setdiff(preds, names(vars)[v])
+
+
+ miceRanger(dpp_data,
+            m = n_impute,
+            maxiter = n_iter,
+            meanMatchCandidates = n_neighbors,
+            vars = vars,
+            num.trees = n_trees,
+            returnModels = TRUE)
+
+}

### 2024-05-02 R/dpp_load.R

@@ -1,0 +1,61 @@
+
+
+#' Load all of the non-form data files
+#'
+#' @author Byron C Jaeger
+#'
+#' @param format used to determine whether you want csv files or sas ones.
+#'   Note that you need to be specific in saying "\\.csv$" or "\\.sas7bdat$".
+#'   If you say something simpler like "csv" the function won't work.
+#'
+#' @return a list of data frames
+dpp_load <- function(format = '\\.csv$'){
+
+ fpath <- file.path(
+  "\\\\medctr.ad.wfubmc.edu",
+  "dfs",
+  "phs_research$",
+  "bancks_2022ada",
+  "DPP",
+  "Data_original",
+  "DPP_Data_2008",
+  "Non-Form_Data",
+  "Data"
+ )
+
+ files <- list.files(fpath, pattern = format)
+ data_names <- str_remove(files, format)
+
+ reader_fun <- switch(format,
+                      "\\.csv$" = readr::read_csv,
+                      "\\.sas7bdat$" = haven::read_sas)
+
+ files_nonform <- file.path(fpath, files) %>%
+  set_names(data_names) %>%
+  map(reader_fun) %>%
+  map(clean_names)
+
+ fpath <- file.path(
+  "\\\\medctr.ad.wfubmc.edu",
+  "dfs",
+  "phs_research$",
+  "bancks_2022ada",
+  "DPP",
+  "Data_original",
+  "DPP_Data_2008",
+  "Form_Data",
+  "Data"
+ )
+
+ files <- list.files(fpath, pattern = format)
+ data_names <- str_remove(files, format)
+
+ files_form <- file.path(fpath, files) %>%
+  set_names(data_names) %>%
+  map(reader_fun) %>%
+  map(clean_names)
+
+ list(non_form = files_nonform,
+      form = files_form)
+
+}

### 2024-05-02 R/dpp_merge.R

@@ -1,0 +1,48 @@
+dpp_merge <- function(dpp_components, daysrand_max = 0) {
+
+ # filter datasets with visits to contain only data at or before baseline
+ for(i in seq_along(dpp_components$non_form)){
+
+  if("visit" %in% names(dpp_components$non_form[[i]])){
+
+   dpp_components$non_form[[i]] <- dpp_components$non_form[[i]] %>%
+    filter(visit %in% c('BAS', 'RUN'), daysrand <= daysrand_max) %>%
+    # NOTE: these ids appear to have duplicated data in the diet file
+    # 10029375 10031210 10033391 10036250
+    # the distinct call here prevents that from causing duplicates to be
+    # brought into the merge statement below
+    distinct()
+
+   # drop these identifiers, merge should only use release_id
+   dpp_components$non_form[[i]]$visit <- NULL
+   dpp_components$non_form[[i]]$daysrand <- NULL
+
+  }
+
+ }
+
+ dpp_nonform <- reduce(dpp_components$non_form, .f = left_join)
+
+ # include forms data (don't worry about filtering to visit/daysrand)
+
+ # education
+ dpp_edu <- dpp_components$form$s06 %>%
+  transmute(
+   release_id,
+   edugrp = case_when(
+    seeduc < 12  ~ "less than high school",
+    seeduc == 12 ~ "high school",
+    seeduc <= 16 ~ "some college to college degree"
+   ),
+   edugrp = factor(edugrp, levels = c("less than high school",
+                                      "high school",
+                                      "some college to college degree"))
+  )
+
+ dpp_nonform %>%
+  left_join(dpp_edu) %>%
+  relocate(edugrp, .before = weight)
+
+
+}
+

### 2024-05-02 R/mesa_bind_rows.R

@@ -1,0 +1,16 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param mesa_visit_one
+#' @param mesa_visit_two
+mesa_bind_rows <- function(mesa_visit_one, mesa_visit_two) {
+
+  bind_rows(mesa_visit_one,
+            mesa_visit_two) %>%
+  arrange(id, visit) %>%
+  group_by(desc(id)) %>%
+  slice(1)
+
+}

### 2024-05-02 R/mesa_load.R

@@ -1,0 +1,89 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param visit
+mesa_load <- function() {
+
+  fpath <- file.path(
+    "\\\\medctr.ad.wfubmc.edu",
+    "dfs",
+    "phs_research$",
+    "bancks_2022ada",
+    "bancks",
+    "MESA",
+    "Analytic data"
+  )
+
+  mesa_visit_one <-
+    read_sas(file.path(fpath, "mesaex1_nodiabv20231128.sas7bdat")) %>%
+    transmute(
+      id = idno,
+      prediabetes = prediabex1,
+      visit = 1L,
+      time_diabetes = exdiabpyr,
+      status_diabetes = incdiab,
+      age_yrs = age1c,
+      sex = factor(female, levels = c(0,1), labels = c("male", "female")),
+      race_ethnicity = factor(race1c,
+                              labels = c("caucasian",
+                                         "african_american",
+                                         "hispanic",
+                                         "other")),
+      education = factor(edugrp,
+                         levels = c(0, 1, 2, 3),
+                         labels = c("less than high school",
+                                    "high school",
+                                    "some college to college degree",
+                                    "some college to college degree")),
+      treatment = factor(NA,
+                         levels = c("lifestyle",
+                                    "metformin",
+                                    "placebo")),
+      glucose_fasting_mgdl = glucos1c,
+      hba1c_percent = NA_real_,
+      chol_ldlv_mgdl = ldl1,
+      chol_trig_mgdl = trig1,
+      bmi = bmi1c
+    )
+
+  mesa_visit_two <-
+    read_sas(file.path(fpath, "mesaex2_nodiab.sas7bdat")) %>%
+    filter() %>%
+    filter() %>%
+    transmute(
+      id = idno,
+      prediabetes = glucos2c >= 100 &
+        glucos2c <= 125 &
+        hba1c2 < 6.5    &
+        hba1c2 > 5.7,
+      visit = 2L,
+      time_diabetes = exdiabpyr26,
+      status_diabetes = incdiab26,
+      age_yrs = age2c,
+      sex = factor(female, levels = c(0,1), labels = c("male", "female")),
+      race_ethnicity = factor(race1c,
+                              labels = c("caucasian",
+                                         "other",
+                                         "african_american",
[…]

### 2024-05-02 _targets.R

@@ -1,0 +1,93 @@
+
+source("packages.R")
+source("conflicts.R")
+
+## Load your R files
+lapply(list.files("./R", full.names = TRUE), source)
+
+library(future)
+library(future.callr)
+plan(callr)
+
+
+manuscript_version <- 1
+
+if(!dir.exists(glue("doc/manuscript-v{manuscript_version}"))){
+  dir.create(glue("doc/manuscript-v{manuscript_version}"))
+}
+
+
+labels_tar <- tar_target(
+  labels,
+  list(variables = c(age_yrs = "Age, years",
+                     race_ethnicity = "Race/ethnicity",
+                     sex = "Sex",
+                     treatment = "Treatment group",
+                     glucose_fasting_mgdl = "Fasting glucose, mg/dl",
+                     hba1c_percent = "HbA1c, %",
+                     chol_trig_mgdl = "Triglycerides, mg/dl",
+                     bmi = "Body mass index, kg/m2"))
+)
+
+dpp_init_tar <- tar_target(dpp_init, load_dpp())
+
+manuscript_tar <- tar_render(
+  manuscript,
+  path = here::here("doc/manuscript.Rmd"),
+  output_file = paste0("manuscript", "-v", manuscript_version, "/",
+                       "manuscript-", basename(here()),
+                       "-v", manuscript_version,
+                       ".docx")
+)
+
+
+# step 1: load all the components of DPP
+dpp_components_tar <- tar_target(
+  dpp_components,
+  dpp_load(format = "\\.sas7bdat$")
+)
+
+# step 2: merge, keeping data collected PRIOR to randomization
+dpp_merged_tar <- tar_target(
+  dpp_merged,
+  dpp_merge(dpp_components)
+)
+
+# step 3, rename variables and derive some
+dpp_formatted_tar <- tar_target(
+  dpp_formatted,
+  dpp_format(dpp_merged)
+)
+
+dpp_excluded_tar <- tar_target(
+  dpp_excluded,
+  dpp_exclude(dpp_formatted)
+)
+
+mesa_init_tar <- tar_target(
+  mesa_init,
+  mesa_load()
+)
+
+mesa_excluded_tar <- tar_target(
+  mesa_excluded,
+  mesa_exclude()
+)
+
+
+
+targets <- list(
+  labels_tar,
+  dpp_components_tar,
+  dpp_merged_tar,
+  dpp_formatted_tar,
+  dpp_excluded_tar,
[…]

### 2024-05-02 conflicts.R

@@ -1,0 +1,5 @@
+conflicts_prefer(dplyr::select)
+conflicts_prefer(dplyr::filter)
+conflicts_prefer(dplyr::slice)
+conflicts_prefer(purrr::set_names)
+

### 2024-05-02 packages.R

@@ -1,0 +1,34 @@
+
+library(conflicted)
+library(dotenv)
+library(targets)
+library(tarchetypes)
+library(aorsf)
+library(tidyverse)
+library(tidymodels)
+library(glue)
+library(janitor)
+library(haven)
+library(grf)
+library(magrittr)
+
+library(survival)
+library(survival.calib)
+library(riskRegression)
+library(ranger)
+library(randomForestSRC)
+library(pec)
+library(glmnet)
+library(party)
+library(xgboost)
+library(miceRanger)
+library(haven)
+
+library(gt)
+library(table.glue)
+library(data.table)
+library(flextable)
+
+# standardizes starting location for filepaths
+# (quite important for docs)
+library(here)

### 2024-05-12 doc/manuscript.Rmd
more lightweight analysis

@@
 
-```
-
-Your name here, your degree,^1^ 
-Second in Command, MS,^1^ 
-Minor Contributor, PhD,^2^ and 
-The Mastermind, PhD^1^
-
-^1^Your department, University, City, State. `r single_par`
-^2^Department of Win, Overtly Pompous University, Irvine, CA. `r single_par`
-
-<br>
-<br>
-<br>
-<br>
-<br>
-<br>
-<br>
-<br>
-<br>
-<br>
-
-**Correspondence to:** \
-Your name, Degree \
-Your Department \
-Additional departmental details if needed \
-Your university \
-Address line 1 \
-Address line 2 \
-Address line 3 \
-your phone number \
-your email `r single_par`
-
-**Word Count:** Abstract: ; Manuscript:
-
-```{r data-prep, include=FALSE}
-
-library(palmerpenguins)
-
-data <- penguins
-
-```
-
-\newpage
-
-# ABSTRACT
-
-\newpage
-
-# INTRODUCTION
-
-Here's how to cite one thing.[@r_language]
-
-Here's how to cite two or more things.[@ambrosius_design_2014; @sprint_research_group_randomized_2015]
-
-# METHODS
-
-# RESULTS
-
-```{r results-prep}
-
-n_penguins <- nrow(data)
-
-mean_bill_length <- mean(data$bill_length_mm, na.rm = TRUE)
-
-mean_mass <- mean(data$body_mass_g, na.rm = TRUE)
-sd_mass <- sd(data$body_mass_g, na.rm = TRUE)
-
-# make a short name for table.glue::table_value - this makes
-# the inline reporting of results less complicated to read.
-tv <- table.glue::table_value
-tg <- table.glue::table_glue
-
-```
-
-Here's how to put some results into the text: A total of `r n_penguins` were included in the current analysis.
-The mean bill length was `r tv(mean_bill_length)`.
-The mean (standard deviation) body mass was `r tg("{mean_mass} ({sd_mass})")`
-
-# DISCUSSION
-
-\newpage
-
-# REFERENCES
-
-<div id="refs"></div>
-
-\newpage
-
-# ACKNOWLEDGMENTS
-
-# FINANCIAL DISCLOSURE
-
-```{r tbl_footers}
-
-ckd_defn <- as_paragraph("Chronic Kidney Disease was defined by an estimated glomerular filtration rate <60 ml/min/1.73 m", as_sup(2L), " based on the 2021 CKD-EPI creatinine equation.")
+withr::with_dir(
+  here::here(),
+  code = {
+    targets::tar_load(c(labels,
+                        analysis,
+                        tbl_exclusion,
+                        tbl_characteristics,
+                        tbl_cuminc))
+  }
+)
 
@@
   IQR  = 'interquartile range',
-  MoCA = 'Montreal cognitive assessment',
-  EHR  = 'electronic health records',
-  CVD  = 'cardiovascular disease'
+  CVD  = 'cardiovascular disease',
+  DPP  = "Diabetes Prevention Program",
+  MESA = "Multi-Ethnic Study of Atherosclerosis"
 )
@@
 
-```
+fn_mesa_baseline <- as_paragraph("Baseline for Multi-Ethnic Study of Atherosclerosis participants was visit 1 or 2 dependent on pre-diabetes status")
 
-**Table 1: Penguin characteristics**
-
-```{r}
-
-tbl_summary(data, by = 'island') %>% 
- as_flex_table() %>% 
- theme_box() %>% 
- align(j = c(2,3,4), align = 'center')
+ft_tidy <- function(x){
+  theme_box(x) %>% 
+    align(align = 'center', part = 'all') %>%
+    align(j = 1, align = 'left', part = 'all')
+}
 
@@
 
-**Table 2: A more customized table**
+Table 0: Participants included.
+
+```{r tbl_exclusion}
+
+tbl_exclusion %>% 
+  flextable() %>% 
+  width(j = 'name', width = 3) %>% 
+  add_footer_lines(
+    values = abbrvs_write(abbr = abbrvs[c("DPP", "MESA")])
+  ) %>% 
+  set_header_labels(name = "Inclusion criteria") %>% 
+  footnote(part = 'body', 
+           i = 2, 
+           j = 'name', 
+           ref_symbols = '1', 
+           value = fn_mesa_baseline) %>% 
+  ft_tidy()
+
+```
+
+\newpage
+
+Table 1: Descriptive table of Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis participants included in the current analysis
 
@@
 
-data %>% 
- group_by(species) %>% 
- slice(1:3) %>% 
- select(-body_mass_g, -sex, -year) %>% 
- mutate(across(where(is.numeric), table_value)) %>% 
- as_grouped_data(groups = 'species') %>% 
- as_flextable(hide_grouplabel = TRUE) %>% 
- set_header_labels(bill_length_mm = 'Bill length, mm',
-                   bill_depth_mm = 'Bill depth, mm',
-                   flipper_length_mm = 'Flipper length, mm') %>% 
- add_header_row(values = c(" ", " Penguin things"),
-                colwidths = c(1, 3)) %>% 
- bg(i = ~!is.na(species), bg = 'grey90') %>% 
- width(width = 1.5) %>% 
- width(j = 1, width = 2) %>% 
- align(part = 'all', j = c(2,3,4), align = 'center') %>% 
- footnote(i = 2, j = 2, 
-          part = 'header',
-          value = as_paragraph("Bill length doesn't correlate with what you think it does."),
-          ref_symbols = '1') %>% 
- footnote(i = 1, j = 1, 
-          part = 'body',
-          value = as_paragraph('Adelies are the best and they know it'), 
-          ref_symbols = '2') %>% 
- add_footer_lines(values = abbrvs_write(abbrvs[c('CI', 'EHR', 'SD')])) %>% 
- add_footer_lines(values = "none of those abbreviations are in the table but they demonstrate how to include abbreviations using my helper functions")
+tbl_characteristics$included %>% 
+  as_flex_table() %>% 
+  ft_tidy()
 
@@
 
-<!-- new page not needed at the end of a block section -->
+\newpage
+
+Table 2: Descriptive table of Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis participants excluded from the current analysis
 
@@
 
-margins_normal <- page_mar(bottom = 2.54 * 0.3937, 
-                           top    = 2.54 * 0.3937, 
-                           right  = 2.54 * 0.3937, 
-                           left   = 2.54 * 0.3937,
-                           gutter = 0)
-
-block_section(
-  prop_section(
-    page_size = page_size(orient = "portrait"),
-    type = "continuous",
-    page_margins = margins_normal
-  )
-)
+tbl_characteristics$excluded %>% 
+  as_flex_table() %>% 
+  ft_tidy()
 
@@
 
-To make a landscape section, use the block_section code here. Note that the section will be applied to all content prior to the block reaching up to the most recent block.
+\newpage
 
-```{r echo = TRUE}
-block_section(
-  prop_section(
-    page_size = page_size(orient = "landscape"),
-    page_margins = page_mar(bottom = 1/8, top = 1/8, 
-                            right = 1/4, left = 1/4),
-    type = "continuous"
-  )
-)
+Table 3. Cumulative incidence of diabetes over follow-up in the Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis
+
+```{r}
+
+tbl_cuminc %>% 
+  ft_tidy() %>% 
+  bg(i = ~ !is.na(group), part = 'body', bg = 'grey80')
+
 ```
 
-To switch back to regular portrait section, add another portrait block. Notice how the block below acts on all content above it up to the prior block.
+\newpage
 
-```{r echo = TRUE}
-block_section(
-  prop_section(
-    page_size = page_size(orient = "portrait"),
-    type = "continuous",
-    page_margins = margins_normal
+Table 4.
+
+```{r}
+
+analysis$eval %>% 
+  pivot_wider(names_from = evaluation, 
+              values_from = c(standard, individualized)) %>% 
+  select(stat, 
+         standard_internal,
+         individualized_internal,
+         standard_external,
+         individualized_external) %>% 
+  separate(stat, into = c("group", "stat"), sep = '\\.\\.', fill = 'left') %>% 
+  mutate(
+    stat = recode(
+      stat,
+      demo_parity = "Demographic parity",
+      equal_oppo = "Equal opportunity",
+      equal_odds = "Equal odds"
+    ),
+    group = recode(
+      group,
+      race = "Race/ethnicity",
+      sex = "Sex"
     )
-)
+  ) %>% 
+  as_grouped_data(group = 'group') %>% 
+  janitor::remove_empty('rows') %>% 
+  as_flextable(hide_grouplabel = TRUE) %>%
+  set_header_labels(
+    stat = "Evaluation statistic",
+    standard_internal = "Standard",
+    individualized_internal = "Individualized",
+    standard_external = "Standard",
+    individualized_external = "Individualized"
+  ) %>% 
+  add_header_row(values = c("Evaluation statistic", 
+                            "Internal Validation", 
+                            "External Validation"), 
+                 colwidths = c(1, 2, 2)) %>% 
+  ft_tidy() %>% 
+  width(width = 1.5) %>% 
+  merge_v(part = 'header', j = 1) %>% 
+  bg(i = ~ !is.na(group), part = 'body', bg = 'grey80')
+
 ```

### 2024-05-12 R/dpp_exclude.R

@@ -24,8 +24,7 @@ dpp_exclude <- function(dpp_formatted) {
  n_participants <- c(n_participants, nrow(e2))
 
  exclusions <- enframe(n_participants) %>%
-  mutate(dropped = abs(c(0, diff(value))),
-         name = c("Study participants",
+  mutate(name = c("Study participants",
                    "Pre-diabetic at baseline",
                    "Randomized to placebo, metformin, or lifestyle"))
 

### 2024-05-12 R/mesa_bind_rows.R

@@ -1,16 +1,0 @@
-#' .. content for \description{} (no empty lines) ..
-#'
-#' .. content for \details{} ..
-#'
-#' @title
-#' @param mesa_visit_one
-#' @param mesa_visit_two
-mesa_bind_rows <- function(mesa_visit_one, mesa_visit_two) {
-
-  bind_rows(mesa_visit_one,
-            mesa_visit_two) %>%
-  arrange(id, visit) %>%
-  group_by(desc(id)) %>%
-  slice(1)
-
-}

### 2024-05-12 R/mesa_exclude.R

@@ -1,0 +1,34 @@
+
+#' @author bcjaeger
+
+mesa_exclude <- function(mesa_init) {
+
+  # don't use nrow because the mesa data were pre-filtered
+  n_participants <- 6814
+
+  e1 <- mesa_init %>%
+      # drop those without pre-diabetes
+    filter(prediabetes == 1) %>%
+    group_by(id) %>%
+    slice(1) %>%
+    ungroup()
+
+  n_participants <- c(n_participants, nrow(e1), NA_integer_)
+
+  exclusions <- enframe(n_participants) %>%
+    mutate(name = c("Study participants",
+                    "Pre-diabetic at baseline",
+                    "Randomized to placebo, metformin, or lifestyle"))
+
+
+  out <- list(included = e1,
+              excluded = setdiff(mesa_init, e1))
+
+  # assert no loss of participants
+  stopifnot(nrow(bind_rows(out)) == nrow(mesa_init))
+
+  attr(out, 'exclusions') <- exclusions
+
+  out
+
+}

### 2024-05-12 R/mesa_load.R

@@ -24,6 +24,8 @@ mesa_load <- function() {
       visit = 1L,
       time_diabetes = exdiabpyr,
       status_diabetes = incdiab,
+      time_death = deathpyr,
+      status_death = death2019,
       age_yrs = age1c,
       sex = factor(female, levels = c(0,1), labels = c("male", "female")),
       race_ethnicity = factor(race1c,
@@ -50,8 +52,6 @@ mesa_load <- function() {
 
   mesa_visit_two <-
     read_sas(file.path(fpath, "mesaex2_nodiab.sas7bdat")) %>%
-    filter() %>%
-    filter() %>%
     transmute(
       id = idno,
       prediabetes = glucos2c >= 100 &
@@ -61,6 +61,8 @@ mesa_load <- function() {
       visit = 2L,
       time_diabetes = exdiabpyr26,
       status_diabetes = incdiab26,
+      time_death = deathpyr26,
+      status_death = death2019,
       age_yrs = age2c,
       sex = factor(female, levels = c(0,1), labels = c("male", "female")),
       race_ethnicity = factor(race1c,
@@ -82,8 +84,6 @@ mesa_load <- function() {
   bind_rows(mesa_visit_one,
             mesa_visit_two) %>%
     drop_na(time_diabetes, status_diabetes) %>%
-    arrange(id, visit) %>%
-    group_by(id) %>%
-    slice(1) %>%
+    arrange(id, visit)
 
 }

### 2024-05-12 R/tabulate_characteristics.R

@@ -1,0 +1,83 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param dpp_excluded
+#' @param mesa_excluded
+tabulate_characteristics <- function(dpp_excluded, mesa_excluded, labels) {
+
+  map2(
+    .x = dpp_excluded,
+    .y = mesa_excluded,
+    .f = tabulate_characteristics_worker,
+    labels = labels
+  )
+
+}
+
+tabulate_characteristics_worker <- function(dpp_data, mesa_data, labels){
+
+  tbl_data <- bind_rows(DPP = dpp_data,
+                        MESA = mesa_data,
+                        .id = 'study') %>%
+    select(
+      study,
+      age_yrs,
+      sex,
+      race_ethnicity,
+      education,
+      glucose_fasting_mgdl,
+      hba1c_percent,
+      homa_ir,
+      homa_beta,
+      bmi,
+      chol_trig_mgdl,
+      chol_ldlc_mgdl,
+      chol_hdl_mgdl
+    )
+
+  cat_vars <- labels$levels
+
+  for(i in names(cat_vars)){
+
+    tbl_data[[i]] <- factor(tbl_data[[i]],
+                            levels = names(labels$levels[[i]]),
+                            labels = labels$levels[[i]])
+
+  }
+
+  .label <- labels$variables
+
+  ctns_vars <- labels$units
+
+  for(i in names(ctns_vars)){
+
+    .label[[i]] <- paste(.label[[i]], labels$units[[i]], sep = ", ")
+
+  }
+
+  out <- tbl_summary(data = tbl_data,
+                     label = .label,
+                     by = 'study',
+                     missing = 'no',
+                     statistic = list(age_yrs = "{mean} ({sd})",
+                                      glucose_fasting_mgdl = "{mean} ({sd})",
+                                      hba1c_percent = "{mean} ({sd})",
+                                      bmi = "{mean} ({sd})",
+                                      homa_ir = "{median} ({p25}, {p75})",
+                                      homa_beta = "{median} ({p25}, {p75})",
+                                      chol_trig_mgdl = "{median} ({p25}, {p75})",
+                                      chol_ldlc_mgdl = "{mean} ({sd})",
+                                      chol_hdl_mgdl = "{mean} ({sd})")) %>%
+    modify_footnote(c(all_stat_cols()) ~ NA)
+
[…]

### 2024-05-12 R/tabulate_cuminc.R

@@ -1,0 +1,103 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param data_analysis
+tabulate_cuminc <- function(data_analysis, pred_horizon = 3) {
+
+  data_cr <- data_analysis %>%
+    mutate(
+      status = if_else(
+        time_death < time_diabetes & status_death == 1,
+        2,
+        status_diabetes
+      ),
+      time = if_else(
+        time_death < time_diabetes & status_death == 1,
+        time_death,
+        time_diabetes
+      ),
+      .status = if_else(time > pred_horizon, 0, status),
+      .time = pmin(pred_horizon, time)
+    ) %>%
+    mutate(status = factor(status)) %>%
+    select(study, time, status, .time, .status)
+
+  # there were no all-cause mortality events, but the code above
+  # would catch them if they were there.
+
+  cuminc_object <- cuminc(ftime = data_cr$time,
+                          fstatus = data_cr$status,
+                          group = data_cr$study)
+
+  cuminc_stats <- cuminc_object[c("DPP 1", "MESA 1")] %>%
+    map(as_tibble) %>%
+    bind_rows(.id = 'study') %>%
+    mutate(timediff = abs(time - 3)) %>%
+    arrange(timediff) %>%
+    group_by(study) %>%
+    slice(1) %>%
+    transmute(
+      study = str_remove(study, " 1$"),
+      cuminc = table_glue(
+        "{100 * est} ({100 * (est-sqrt(var)*1.96)}, {100 * (est+sqrt(var)*1.96)})"
+      )
+    )
+
+  sf <- survfit(Surv(time, status==0)~study, data= data_cr)
+
+  sf_stats <- survival:::survmean(sf, rmean = 'none')$matrix %>%
+    as_tibble(rownames = 'study') %>%
+    mutate(study = str_remove(study, "^study=")) %>%
+    transmute(
+      study,
+      fup_median = table_glue(
+        "{median} ({`0.95LCL`}, {`0.95UCL`})"
+      )
+    )
+
+  data_tbl <- data_cr %>%
+    group_by(study) %>%
+    summarize(n_obs = n(),
+              n_events = sum(.status==1),
+              inc_100 = 100 * sum(.status==1) / sum(.time),
+              fup_total = sum(time)) %>%
+    left_join(sf_stats) %>%
+    left_join(cuminc_stats) %>%
+    mutate(across(where(is.numeric), table_value)) %>%
[…]

### 2024-05-12 R/tabulate_exclusions.R

@@ -1,0 +1,16 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param dpp_excluded
+#' @param mesa_excluded
+tabulate_exclusions <- function(dpp_excluded, mesa_excluded) {
+
+  bind_rows(DPP = attr(dpp_excluded, 'exclusions'),
+            MESA = attr(mesa_excluded, 'exclusions'),
+            .id = 'study') %>%
+    mutate(value = table_value(as.integer(value))) %>%
+    pivot_wider(names_from = study, values_from = value)
+
+}

### 2024-05-12 R/validate_preds.R

@@ -1,0 +1,365 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param data_analysis
+validate_preds <- function(data_analysis,
+                           pred_horizon = 3,
+                           risk_cut = 0.125) {
+
+  initial_split <- data_analysis %>%
+    mutate(surv = Surv(time_diabetes, status_diabetes == 1)) %>%
+    select(
+      study,
+      surv,
+      age_yrs,
+      race_ethnicity,
+      sex,
+      treatment,
+      glucose_fasting_mgdl,
+      hba1c_percent,
+      chol_trig_mgdl,
+      bmi
+    ) %>%
+    split(.$study) %>%
+    map(select, -study)
+
+  derivation <- initial_split$DPP
+  validation <- initial_split$MESA
+
+  derivation_folds <- vfold_cv(derivation)
+
+  rec_standard <- recipe(surv ~ ., data = derivation) %>%
+    step_impute_mean(all_numeric_predictors()) %>%
+    step_impute_mode(all_nominal_predictors()) %>%
+    update_role(race_ethnicity, new_role = 'id')
+
+  rec_interaction <- rec_standard %>%
+    step_interact(terms = ~treatment:matches("age_yrs|bmi|^glucose"))
+
+  cph_spec <- proportional_hazards() %>%
+    set_engine("survival") %>%
+    set_mode("censored regression")
+
+  wf_standard <- workflow() %>%
+    add_recipe(rec_standard) %>%
+    add_model(cph_spec)
+
+  wf_individualized <- workflow() %>%
+    add_recipe(rec_interaction) %>%
+    add_model(cph_spec)
+
+  wf <- list(
+    standard = wf_standard,
+    individualized = wf_individualized
+  )
+
+  # internal validation ----
+
+  preds_out <- list()
+  test_out <- list()
+
+  for(i in seq(nrow(derivation_folds))){
+
+    train <- training(derivation_folds$splits[[i]])
+    test <- testing(derivation_folds$splits[[i]])
+
+    fits <- map(wf, .f = fit, data = train)
+
+    preds_out[[i]] <- map_dfr(fits,
+                              .f = predict,
+                              .id = 'model',
+                              new_data = test,
+                              eval_time = pred_horizon,
+                              type = 'survival') %>%
+      group_by(model) %>%
+      mutate(.id = seq(n()))
+
+    test_out[[i]] <- tibble(time = test$surv[,1],
[…]

### 2024-05-12 R/visualize_cuminc.R

@@ -1,0 +1,57 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param data_analysis
+visualize_cuminc <- function(data_analysis) {
+
+  data_cr <- data_analysis %>%
+    mutate(
+      status = if_else(
+        time_death < time_diabetes & status_death == 1,
+        2,
+        status_diabetes
+      ),
+      time = if_else(
+        time_death < time_diabetes & status_death == 1,
+        time_death,
+        time_diabetes
+      )
+    ) %>%
+    mutate(status = factor(status)) %>%
+    select(study, time, status)
+
+  cuminc_object <- cuminc(ftime = data_cr$time,
+                          fstatus = data_cr$status,
+                          group = data_cr$study)
+
+  pval <- table.glue::table_pvalue(cuminc_object$Tests[,'pv'])
+
+  data_gg <- cuminc_object[c("DPP 1", "MESA 1")] %>%
+    map(as_tibble) %>%
+    bind_rows(.id = 'study') %>%
+    mutate(study = str_remove(study, " 1$")) %>%
+    filter(time < 5)
+
+  ggplot(data_gg) +
+    aes(x = time, y = est) +
+    geom_line(aes(color = study)) +
+    theme_bw() +
+    theme(panel.grid = element_blank(),
+          legend.position = 'right',
+          legend.justification = 'top') +
+    scale_y_continuous(labels = scales::percent) +
+    labs(x = "Time, years since baseline",
+         y = "Cumulative incidence, type 2 diabetes",
+         color = "") +
+    annotate(
+      geom = 'text',
+      x = 4,
+      y = 0.1,
+      label = glue(
+        "Gray's test for equivalent\ncumulative incidence: P = {pval}"
+      )
+    )
+
+}

### 2024-05-12 _targets.R

@@ -19,27 +19,51 @@ if(!dir.exists(glue("doc/manuscript-v{manuscript_version}"))){
 
 labels_tar <- tar_target(
   labels,
-  list(variables = c(age_yrs = "Age, years",
-                     race_ethnicity = "Race/ethnicity",
-                     sex = "Sex",
-                     treatment = "Treatment group",
-                     glucose_fasting_mgdl = "Fasting glucose, mg/dl",
-                     hba1c_percent = "HbA1c, %",
-                     chol_trig_mgdl = "Triglycerides, mg/dl",
-                     bmi = "Body mass index, kg/m2"))
+  list(
+    variables = c(
+      age_yrs = "Age",
+      race_ethnicity = "Race/ethnicity",
+      sex = "Sex",
+      education = "Educational attainment",
+      treatment = "Treatment group",
+      glucose_fasting_mgdl = "Fasting glucose",
+      homa_ir = "HOMA-Insulin Resistance",
+      homa_beta = "HOMA-Beta cell function",
+      hba1c_percent = "Glycated Hemoglobin",
+      chol_trig_mgdl = "Triglycerides",
+      chol_ldlc_mgdl = "Low-density lipoprotein cholesterol",
+      chol_hdl_mgdl = "High-density lipoprotein cholesterol",
+      bmi = "Body mass index"),
+    units = c(
+      age_yrs = "years",
+      glucose_fasting_mgdl = "mg/dl",
+      hba1c_percent = "%",
+      chol_trig_mgdl = "mg/dl",
+      bmi = "kg/m2",
+      chol_ldlc_mgdl = "mg/dl",
+      chol_hdl_mgdl = "mg/dl"
+    ),
+    levels = list(
+      sex = c(
+        "male" = "Male",
+        "female" = "Female"
+      ),
+      race_ethnicity = c(
+        "caucasian" = "Non-Hispanic White",
+        "african_american" = "Non-Hispanic Black",
+        "hispanic" = "Hispanic",
+        "other" = "Other/Chinese"
+      ),
+      education = c(
+        "less than high school" = "< High School",
+        "high school" = "High School Graduate",
+        "some college to college degree" = "Some College or College Graduate"
+      )
+    )
+  )
 )
 
 dpp_init_tar <- tar_target(dpp_init, load_dpp())
-
[…]

### 2024-05-12 conflicts.R

@@ -2,4 +2,7 @@ conflicts_prefer(dplyr::select)
 conflicts_prefer(dplyr::filter)
 conflicts_prefer(dplyr::slice)
 conflicts_prefer(purrr::set_names)
+conflicts_prefer(cmprsk::cuminc)
+conflicts_prefer(flextable::as_flextable)
+conflicts_prefer(parsnip::fit)
 

### 2024-05-12 packages.R

@@ -6,28 +6,27 @@ library(tarchetypes)
 library(aorsf)
 library(tidyverse)
 library(tidymodels)
+library(censored)
 library(glue)
 library(janitor)
 library(haven)
 library(grf)
 library(magrittr)
+library(haven)
 
 library(survival)
 library(survival.calib)
 library(riskRegression)
-library(ranger)
-library(randomForestSRC)
-library(pec)
-library(glmnet)
-library(party)
-library(xgboost)
-library(miceRanger)
-library(haven)
+library(nricens)
+library(dcurves)
 
 library(gt)
+library(gtsummary)
 library(table.glue)
 library(data.table)
 library(flextable)
+library(cmprsk)
+
 
 # standardizes starting location for filepaths
 # (quite important for docs)

### 2025-04-04 doc/review/reviewer_response.Rmd
clean up and add reproduction tools

---
title: "Individualized risk prediction for type 2 diabetes"
bibliography: refs.bib
csl: jama.csl
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
    keep_md: true
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      dpi = 600, 
                      fig.width = 8, 
                      fig.height = 6,
                      fig.path = "doc/fig/")

# load standard packages for RMD word docs
library(officer)
library(flextable)
library(gtsummary)
library(dplyr)
library(purrr)
library(tidyverse)
library(survival)
library(broom)

# specify a rounding specification 
library(table.glue)

conflicted::conflict_prefer('as_flextable', winner = 'flextable')
conflicted::conflict_prefer('filter', winner = 'dplyr')

rspec <- round_spec() |>
  round_using_magnitude(digits = c(2, 1, 0),
                        breaks = c(1, 10, Inf))

# save it to options:
names(rspec) <- paste('table.glue', names(rspec), sep = '.')
options(rspec)

# default is double space 
# single spacing for title page, abstract, captions
single_par <- fp_par(line_spacing=1)

withr::with_dir(
  here::here(),
  code = {
    targets::tar_load(c(labels,
                        reviewer_1_cstats,
                        data_analysis,
                        analysis))
  }
)

margins_normal <- page_mar(bottom = 2.54 * 0.1,  #0.3937, 
                           top    = 2.54 * 0.1,  #0.3937, 
                           right  = 2.54 * 0.1,  #0.3937, 
                           left   = 2.54 * 0.1, #0.3937,
                           gutter = 0)

page_long_above <- function(){

  block_section(
    prop_section(
      page_size = page_size(orient = "portrait"),
      type = "continuous",
      page_margins = margins_normal
    )
  )

}

page_wide_above <- function(){
  block_section(
    prop_section(
      page_size = page_size(orient = "landscape"),
      page_margins = margins_normal,
      type = "continuous"
    )
  )
}

```

```{r tbl_abbrvs}

abbrvs <- list(
  CI   = 'confidence interval',
  FI   = 'frailty index',
  SD   = 'standard deviation',
  IQR  = 'interquartile range',
  CVD  = 'cardiovascular disease',
  DPP  = "Diabetes Prevention Program",
  MESA = "Multi-Ethnic Study of Atherosclerosis",
  NRI  = "Net reclassification index",
  IDI  = "Integrated discrimination index",
  AUC  = "Area underneath the receiver-operator characteristic curve",
  IPA  = "Index of prediction accuracy",
  HOMA = "Homeostatic Model Assessment"
)

abbrvs_paste <- function(strings){

  front <- 'Abbreviations'

  matter <- glue::glue_collapse(strings, sep = '; ', last = '; and ')

  paste(front, matter, sep = ': ')

}

abbrvs_write <- function(abbr){

  sorted <- abbr[sort(names(abbr))]

  strings <- map2_chr(names(sorted), sorted, paste, sep = ' = ')

  as_paragraph(abbrvs_paste(strings))

}

fn_mesa_baseline <- as_paragraph("Baseline for Multi-Ethnic Study of Atherosclerosis participants was visit 1 or 2 dependent on pre-diabetes status")

ft_tidy <- function(x){
  theme_box(x) %>% 
    align(align = 'center', part = 'all') %>%
    align(j = 1, align = 'left', part = 'all')
}

```

C-statistics (95% confidence interval) were `r reviewer_1_cstats$data_internal$glucose_fasting_mgdl` and 
`r reviewer_1_cstats$data_internal$hba1c_percent` for FPG and Hba1c, respectively, for internal validation. In external validation, the respective C-statistics were `r reviewer_1_cstats$data_external$glucose_fasting_mgdl` and 
`r reviewer_1_cstats$data_external$hba1c_percent`. Noting that the combination of FPG and Hba1c consistently obtained higher C-statistic and index of prediction accuracy compared to the individual variables, we replicated our main analysis with an additional model that was restricted to these two predictors (Table R1)

`r page_long_above()`

Table R1. Comparison of the individualized model with standard model.

```{r}

analysis$eval %>% 
  drop_na(reviewer_1) %>% 
  pivot_wider(names_from = evaluation, 
              values_from = c(standard, reviewer_1, individualized)) %>% 
  select(stat, 
         subgroup,
         reviewer_1_internal,
         standard_internal,
         individualized_internal,
         reviewer_1_external,
         standard_external,
         individualized_external) %>% 
  as_grouped_data('subgroup') %>% 
  as_flextable(hide_grouplabel = TRUE) %>%
  set_header_labels(
    stat = "Evaluation statistic",
    standard_internal = "Standard",
    reviewer_1_internal = "FPG + Hba1c",
    individualized_internal = "Individualized",
    standard_external = "Standard",
    reviewer_1_external = "FPG + Hba1c",
    individualized_external = "Individualized"
  ) %>% 
  add_header_row(values = c("Evaluation statistic", 
                            "Internal Validation", 
                            "External Validation"), 
                 colwidths = c(1, 3, 3)) %>% 
  ft_tidy() %>% 
  width(width = 1.5) %>% 
  merge_v(part = 'header', j = 1) %>% 
  add_footer_lines(
    values = abbrvs_write(abbr = abbrvs[c("AUC", "IPA")])
  ) %>% 
  footnote(
    i = 1,
    j = 1,
    part = 'header',
    value = as_paragraph('Table values are scaled by a factor of 100 for ease of interpretation.'),
    ref_symbols = '1'
  ) %>% 
  footnote(
    i = 1,
    j = 2,
    part = 'header',
    value = as_paragraph('Internal validation results are based on 10-fold cross-validation in the Diabetes Prevention Program data'),
    ref_symbols = '2'
  ) %>% 
  footnote(
    i = 1,
    j = 5,
    part = 'header',
    value = as_paragraph('External validation results are based on application of models fitted to the Diabetes Prevention Program data to the Multi-Ethnic Study of Atherosclerosis data.'),
    ref_symbols = '3'
  ) 
```

`r page_wide_above()`

```{r}

data_dpp <- data_analysis %>%
  filter(study == "DPP") %>% 
  mutate(age_cat = cut(age_yrs, breaks = c(0, 55, 65, Inf)))

fit_cph <- 
  coxph(Surv(time_diabetes, status_diabetes) ~ sex + education +
          chol_trig_mgdl + bmi + hba1c_percent + glucose_fasting_mgdl +
          treatment*age_yrs, data = data_dpp)

pvals_multivariable <- tidy(anova(fit_cph)) %>% 
  select(term, p.value) %>% 
  filter(str_detect(term, 'age_yrs')) %>% 
  mutate(term = c("main", "tx_interaction"),
         p.value = table_pvalue(p.value, drop_leading_zero = F)) %>% 
  deframe()

data_inc <- cmprsk::cuminc(ftime = data_dpp$time_diabetes,
                           fstatus = data_dpp$status_diabetes,
                           group = data_dpp$age_cat)

pval_unadjust <- data_inc$Tests %>% 
  as_tibble() %>% 
  mutate(pv = table_pvalue(pv, drop_leading_zero = FALSE))

data_gg <- data_inc
data_gg$Tests <- NULL

data_gg %<>%
  as.list() %>%
  map_dfr(~tibble(time = .x$time, est = .x$est), .id = 'age_group') %>%
  filter(time < 3) %>% 
  mutate(age_group = factor(age_group, 
                            levels = c("(0,55] 1",
                                       "(55,65] 1",
                                       "(65,Inf] 1"),
                            labels = c("< 55 years",
                                       "55 to 65 years",
                                       "> 65 years")))

```

We examined the cumulative incidence of diabetes by age groups (age < 55 years, age 55 to 65 years, and age > 65 years) in the DPP. We did not detect a difference in the cumulative incidence curves (Figure R1; p = `r pval_unadjust$pv`). We also did not detect an assocation between continuous age and incident diagnosis of diabetes in a Cox regression model with adjustment for sex, eduction, triglycerides, BMI, Hba1c, fasting plasma glucose, and randomized treatment (p = `r pvals_multivariable['main']`).

\newpage

Figure R1: Cumulative incidence of diabetes diagnosis in the DPP by age group.

```{r}

ggplot(data_gg, aes(x=time, y=est, color = age_group)) +
  geom_line(linewidth = 1) + 
  theme_bw() + 
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Time, years", 
       y = "Cumulative incidence of diabetes, %",
       color = "Age group")

```

`r page_long_above()`

### 2025-04-04 README.Rmd
clean up and add reproduction tools

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# DPP-individualized-risk-prediction

<!-- badges: start -->
<!-- badges: end -->

## Reproducing

We use `targets` to coordinate this project. Results can be reproduced by opening the project and:

1. install dependencies by running `renv::restore()` in your R console.

2. run the pipeline by running `targets::tar_make()`.

### 2025-04-04 doc/manuscript.Rmd
clean up and add reproduction tools

@@
 ---
-title: "Superior paper quality with R-Markdown: The OBVIOUS Study"
+title: "Individualized risk prediction for type 2 diabetes"
 bibliography: refs.bib
@@
                       fig.width = 11, 
-                      fig.height = 6)
+                      fig.height = 6,
+                      fig.path = "fig/")
 
@@
 library(purrr)
+library(tidyverse)
 
@@
                         tbl_characteristics,
-                        tbl_cuminc))
+                        tbl_reclass,
+                        tbl_cph,
+                        tbl_optim,
+                        tbl_cuminc,
+                        tbl_characteristics_optim,
+                        fig_dcurve,
+                        fig_calib_internal,
+                        fig_calib_external))
   }
@@
   DPP  = "Diabetes Prevention Program",
-  MESA = "Multi-Ethnic Study of Atherosclerosis"
+  MESA = "Multi-Ethnic Study of Atherosclerosis",
+  NRI  = "Net reclassification index",
+  IDI  = "Integrated discrimination index",
+  AUC  = "Area underneath the receiver-operator characteristic curve",
+  IPA  = "Index of prediction accuracy",
+  HOMA = "Homeostatic Model Assessment"
 )
@@
 
+New result: The number needed to treat to prevent 1 diabetes case within the next three years:\
+
+Using model recommendation: `r tbl_characteristics_optim$nnt$nnt[1]`\
+Using lifestyle for everyone: `r tbl_characteristics_optim$nnt$nnt[2]`\
+Using metformin for everyone: `r tbl_characteristics_optim$nnt$nnt[3]`\
+
+Discrimination was estimated by concordance (C-) statistic, which measures the probability of assigning higher risk to a case versus a non-case.
+Calibration was evaluated in terms of agreement between average observed risk and deciles of predicted risk.
+The index of prediction accuracy (IPA) combines discrimination and calibration in one summary value that ranges from 0 (poor) to 1 (perfect).
+Net benefit is based on the weighted difference in the true versus false positive rate of a prediction model at a given decision threshold, i.e., $$\text{net benefit} = \text{true positive rate} - (\text{false positive rate} * \text{exchange rate}),$$ where the exchange rate is the odds of the threshold probability used for clinical decision making.
+Net benefit can be interpreted as the number of true-positive cases identified per 100 patients screened.
+
 \newpage
@@
   as_flex_table() %>% 
-  ft_tidy()
+  ft_tidy() %>% 
+  add_footer_lines(
+    values = abbrvs_write(abbr = abbrvs[c("DPP", "MESA", "HOMA")])
+  )
 
@@
 
-Table 4.
+Table 4A. Comparison of the individualized model with standard model.
 
@@
 analysis$eval %>% 
+  select(-reviewer_1) %>% 
   pivot_wider(names_from = evaluation, 
               values_from = c(standard, individualized)) %>% 
+  filter(is.na(subgroup) | subgroup == "Overall") %>% 
   select(stat, 
@@
       stat,
+      "NRI" = "NRI, Overall",
+      "NRI+" = "NRI, Positive",
+      "NRI-" = "NRI, Negative",
       demo_parity = "Demographic parity",
@@
   merge_v(part = 'header', j = 1) %>% 
-  bg(i = ~ !is.na(group), part = 'body', bg = 'grey80')
+  bg(i = ~ !is.na(group),
+     part = 'body',
+     bg = 'grey80') %>%
+  padding(i = ~ is.na(group) & !str_detect(stat, "NRI|IDI|AUC|IPA"), 
+          j = 1,
+          part = 'body', 
+          padding.left = 10) %>% 
+  add_footer_lines(
+    values = abbrvs_write(abbr = abbrvs[c("NRI", "AUC", "IPA")])
+  ) %>% 
+  footnote(
+    i = 1,
+    j = 1,
+    part = 'header',
+    value = as_paragraph('Table values are scaled by a factor of 100 for ease of interpretation.'),
+    ref_symbols = '1'
+  ) %>% 
+  footnote(
+    i = 1,
+    j = 2,
+    part = 'header',
+    value = as_paragraph('Internal validation results are based on 10-fold cross-validation in the Diabetes Prevention Program data'),
+    ref_symbols = '2'
+  ) %>% 
+  footnote(
+    i = 1,
+    j = 4,
+    part = 'header',
+    value = as_paragraph('External validation results are based on application of models fitted to the Diabetes Prevention Program data to the Multi-Ethnic Study of Atherosclerosis data.'),
+    ref_symbols = '3'
+  )  %>% 
+  footnote(
+    i = 7, j = 1,
+    part = 'body',
+    ref_symbols = "4",
+    value = as_paragraph(
+      "Demographic parity is satisfied when a model's predictions have the same predicted positive rate across groups."
+    )
+  ) %>% 
+  footnote(
+    i = 8, j = 1,
+    part = 'body',
+    ref_symbols = "5",
+    value = as_paragraph(
+      "Equal opportunity is satisfied when a model's predictions have the same true positive and false negative rates across protected groups."
+    )
+  ) %>% 
+  footnote(
+    i = 9, j = 1,
+    part = 'body',
+    ref_symbols = "6",
+    value = as_paragraph(
+      "Equal odds is satisfied when a model's predictions have the same false positive, true positive, false negative, and true negative rates across protected groups."
+    )
+  )
 
 ```
+
+\newpage
+
+Table 4B: Subgroup comparison of the individualized model with standard model.
+
+```{r}
+
+analysis$eval %>% 
+  pivot_wider(names_from = evaluation, 
+              values_from = c(standard, individualized)) %>% 
+  filter(!is.na(subgroup) & subgroup != "Overall") %>% 
+  select(stat, 
+         subgroup,
+         standard_internal,
+         individualized_internal,
+         standard_external,
+         individualized_external) %>% 
+  as_grouped_data(group = 'subgroup') %>% 
+  janitor::remove_empty('rows') %>% 
+  as_flextable(hide_grouplabel = TRUE) %>%
+  set_header_labels(
+    stat = "Evaluation statistic",
+    standard_internal = "Standard",
+    individualized_internal = "Individualized",
+    standard_external = "Standard",
+    individualized_external = "Individualized"
+  ) %>% 
+  add_header_row(values = c("Evaluation statistic", 
+                            "Internal Validation", 
+                            "External Validation"), 
+                 colwidths = c(1, 2, 2)) %>% 
+  ft_tidy() %>% 
+  width(width = 1.5) %>% 
+  merge_v(part = 'header', j = 1) %>% 
+  bg(i = ~ !is.na(subgroup),
+     part = 'body',
+     bg = 'grey80') %>%
+  padding(i = ~ is.na(subgroup) & !str_detect(stat, "NRI|AUC|IPA"), 
+          j = 1,
+          part = 'body', 
+          padding.left = 10) %>% 
+  add_footer_lines(
+    values = abbrvs_write(abbr = abbrvs[c("AUC", "IPA")])
+  ) %>% 
+  footnote(
+    i = 1,
+    j = 1,
+    part = 'header',
+    value = as_paragraph('Table values are scaled by a factor of 100 for ease of interpretation.'),
+    ref_symbols = '1'
+  ) %>% 
+  footnote(
+    i = 1,
+    j = 2,
+    part = 'header',
+    value = as_paragraph('Internal validation results are based on 10-fold cross-validation in the Diabetes Prevention Program data'),
+    ref_symbols = '2'
+  ) %>% 
+  footnote(
+    i = 1,
+    j = 4,
+    part = 'header',
+    value = as_paragraph('External validation results are based on application of models fitted to the Diabetes Prevention Program data to the Multi-Ethnic Study of Atherosclerosis data.'),
+    ref_symbols = '3'
+  )
+
+```
+
+\newpage
+
+Table 5A: Reclassification matrix comparing individualized versus standard risk classification among participants in the Diabetes Prevention Program (internal) and the Multi-Ethnic Study of Atherosclerosis (external)
+
+```{r}
+
+tbl_reclass$set %>% 
+  mutate(set = paste(str_to_title(set), "data")) %>% 
+  as_grouped_data(groups = 'set') %>% 
+  as_flextable(hide_grouplabel = TRUE) %>% 
+  ft_tidy() %>% 
+  width(j = 1, width = 1.5) %>% 
+  width(j = c(2,3,4), width = 1.2) %>% 
+  bg(i = ~ !is.na(set), part = 'body', bg = 'grey80') %>% 
+  padding(i = ~ is.na(set), part = 'body', padding.left = 10, j=1) %>% 
+  set_header_labels(standard = "Standard risk categories") %>% 
+  add_header_row(values = c("Standard risk categories", 
+                            "Individualized risk categories"), 
+                 colwidths = c(1, 3)) %>% 
+  merge_v(part = 'header')
+
+```
+
+\newpage
+
+Table 5B: Reclassification matrix comparing individualized versus standard risk classification among men and women.
+
+```{r}
+
+tbl_reclass$sex %>% 
+  mutate(sex = recode(sex, 'male' = "Men", "female" = "Women")) %>% 
+  as_grouped_data(groups = 'sex') %>% 
+  as_flextable(hide_grouplabel = TRUE) %>% 
+  ft_tidy() %>% 
+  width(j = 1, width = 1.5) %>% 
+  width(j = c(2,3,4), width = 1.2) %>% 
+  bg(i = ~ !is.na(sex), part = 'body', bg = 'grey80') %>% 
+  padding(i = ~ is.na(sex), part = 'body', padding.left = 10, j=1) %>% 
+  set_header_labels(standard = "Standard risk categories") %>% 
+  add_header_row(values = c("Standard risk categories", 
+                            "Individualized risk categories"), 
+                 colwidths = c(1, 3)) %>% 
+  merge_v(part = 'header')
+
+```
+
+\newpage
+
+Table 5C: Reclassification matrix comparing individualized versus standard risk classification among men and women.
+
+```{r}
+
+tbl_reclass$race %>% 
+  mutate(
+    race_ethnicity = recode(
+      race_ethnicity, 
+      "caucasian" = "Caucasian", 
+      "african_american" = "African American",
+      "hispanic" = "Hispanic",
+      "other" = "Other"
+    )
+  ) %>% 
+  as_grouped_data(groups = 'race_ethnicity') %>% 
+  as_flextable(hide_grouplabel = TRUE) %>% 
+  ft_tidy() %>% 
+  width(j = 1, width = 1.5) %>% 
+  width(j = c(2,3,4), width = 1.2) %>% 
+  bg(i = ~ !is.na(race_ethnicity), part = 'body', bg = 'grey80') %>% 
+  padding(i = ~ is.na(race_ethnicity), part = 'body', padding.left = 10, j=1) %>% 
+  set_header_labels(standard = "Standard risk categories") %>% 
+  add_header_row(values = c("Standard risk categories", 
+                            "Individualized risk categories"), 
+                 colwidths = c(1, 3)) %>% 
+  merge_v(part = 'header')
+
+```
+
+\newpage
+
+Table 6: Optimal preventive intervention and 3-year counterfactual risk for type 2 diabetes from an individualized risk prediction model in the Diabetes Prevention Program
+
+```{r}
+
+tbl_optim %>% 
+  mutate(tx_opt = str_to_title(tx_opt),
+         type = str_to_title(type),
+         type = factor(type, levels = c("Internal", "External"))) %>% 
+  arrange(type) %>% 
+  as_grouped_data(groups = 'type') %>% 
+  as_flextable(hide_grouplabel = TRUE) %>% 
+  ft_tidy() %>% 
+  width(j = 1, width = 1.5) %>% 
+  width(j = c(2,3,4,5), width = 1.2) %>% 
+  bg(i = ~ !is.na(type), part = 'body', bg = 'grey80') %>% 
+  padding(i = ~ is.na(type), part = 'body', padding.left = 10, j=1) %>% 
+  set_header_labels(tx_opt = "Optimal intervention",
+                    tbn = "N (%)",
+                    lifestyle = "Lifestyle",
+                    metformin = "Metformin",
+                    placebo = "Placebo") %>% 
+  add_header_row(
+    values = c("Optimal intervention",
+               "N (%)",
+               "Counterfactual 3-year T2D Predicted Risk, Mean (SD)"), 
+    colwidths = c(1, 1, 3)
+  ) %>% 
+  merge_v(part = 'header')
+
+```
+
+\newpage
+
+Table 7: Summary of and instructions for using the individualized prediction model for type 2 diabetes.
+
+```{r}
+
+tbl_grouped_data <- tbl_cph$tbl_data %>%
+  arrange(!is.na(variable)) %>%
+  as_grouped_data(groups = 'variable') %>%
+  remove_empty('rows')
+
+padding_index <- with(
+  tbl_grouped_data,
+  is.na(variable) & !is.na(level)
+)
+
+# nonsense fix; short on time
+padding_index[1:2] <- FALSE
+
+tbl_grouped_data %>%
+  as_flextable(hide_grouplabel = TRUE)  %>%
+  ft_tidy() %>% 
+  set_header_labels(
+    level = "Variable", tbl_value = "Hazard ratio (95% CI)"
+  ) %>%
+  align(align = 'center', part = 'all') %>%
+  align(j = 1, align = 'left', part = 'all') %>%
+  padding(i = padding_index,
+          j = 1,
+          padding.left = 15) %>%
+  width(j = 1, width = 2) %>%
+  width(j = c(2), width = 2) %>%
+  footnote(
+    i = 1,
+    j = 1,
+    part = 'header',
+    ref_symbols = '1',
+    value = as_paragraph(
+      "Predictor variables included in the table were selected",
+      " apriori based on clinical availability and known associations."
+    )
+  ) %>%
+  footnote(
+    i = 1,
+    j = 2,
+    part = 'header',
+    ref_symbols = '2',
+    value = as_paragraph(
+      paste(
+        "Hazard ratios are adjusted for all variables listed in the table.",
+        "Hazard ratios for continuous variables correspond to",
+        " a one standard deviation change in the variable."
+      )
+    )
+  ) %>% 
+  add_footer_lines(
+    value = as_paragraph(tbl_cph$instructions)
+  )
+
+block_section(
+  prop_section(
+    page_size = page_size(orient = "portrait"),
+    type = "continuous"
+  )
+)
+```
+
+\newpage
+
+Table 8: Characteristics of participants stratified by recommended treatment to prevent diabetes according to the individualized prediction model.
+
+```{r}
+tbl_characteristics_optim$tbl %>% 
+  ft_tidy() %>% 
+  align(align = 'center', part = 'all') %>%
+  align(j = 1, align = 'left', part = 'all') %>% 
+  add_header_row(values = c("Characteristic", 
+                            "Optimal predicted treatment"),
+                 colwidths = c(1, 2)) %>% 
+  merge_v(part = 'header')
+
+```
+
+\newpage
+
+Figure 1: Decision curve analysis of the individualized and standard model in internal and external validation.
+
+```{r fig-1}
+
+library(patchwork)
+
+p1 <- fig_dcurve$data_internal + labs(title="Internal validation")
+p2 <- fig_dcurve$data_external + labs(title="External validation")
+
+(guide_area() / (p1+p2)) + 
+  plot_layout(guides = "collect", 
+              heights = unit(c(1, 1), c("cm", "null"))) & 
+  theme(legend.position = "top")
+
+```
+
+\newpage
+
+Figure 2: Calibration of the individualized model in internal and external validation.
+
+```{r fig-2}
+
+p1 <- fig_calib_internal$overall + 
+  labs(title="Internal validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+p2 <- fig_calib_external$overall + 
+  labs(title="External validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+(p1 + p2) + 
+  plot_layout(guides = "collect", 
+              axes = 'collect')
+
+```
+
+\newpage
+
+Figure 3: Calibration of the individualized model in internal and external validation among male study participants
+
+```{r fig-3}
+
+p1 <- fig_calib_internal$men + 
+  labs(title="Internal validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+p2 <- fig_calib_external$men + 
+  labs(title="External validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+(p1 + p2) + 
+  plot_layout(guides = "collect", 
+              axes = 'collect')
+
+```
+
+\newpage
+
+Figure 4: Calibration of the individualized model in internal and external validation among female study participants
+
+```{r fig-4}
+
+p1 <- fig_calib_internal$women + 
+  labs(title="Internal validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+p2 <- fig_calib_external$women + 
+  labs(title="External validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+(p1 + p2) + 
+  plot_layout(guides = "collect", 
+              axes = 'collect')
+
+```
+
+\newpage
+
+Figure 5: Calibration of the individualized model in internal and external validation among African American study participants
+
+```{r fig-5}
+
+p1 <- fig_calib_internal$black + 
+  labs(title="Internal validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+p2 <- fig_calib_external$black + 
+  labs(title="External validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+(p1 + p2) + 
+  plot_layout(guides = "collect", 
+              axes = 'collect')
+
+```
+
+\newpage
+
+Figure 6A: Calibration of the individualized model in internal and external validation among Caucasian study participants
+
+```{r fig-6a}
+
+p1 <- fig_calib_internal$white + 
+  labs(title="Internal validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+p2 <- fig_calib_external$white + 
+  labs(title="External validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+(p1 + p2) + 
+  plot_layout(guides = "collect", 
+              axes = 'collect')
+
+```
+
+\newpage
+
+Figure 6B: Calibration of the individualized model in internal and external validation among Hispanic study participants
+
+```{r fig-6b}
+
+p1 <- fig_calib_internal$hispanic + 
+  labs(title="Internal validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+p2 <- fig_calib_external$hispanic + 
+  labs(title="External validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+(p1 + p2) + 
+  plot_layout(guides = "collect", 
+              axes = 'collect')
+
+```
+
+\newpage
+
+Figure 6C: Calibration of the individualized model in internal and external validation among study participants who identified their race/ethnicity as a category other than African American, Caucasian, or Hispanic.
+
+```{r fig-6c}
+
+p1 <- fig_calib_internal$other + 
+  labs(title="Internal validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+p2 <- fig_calib_external$other + 
+  labs(title="External validation",
+       x = "Predicted 3-year risk, %",
+       y = "Observed 3-year risk, %")
+
+(p1 + p2) + 
+  plot_layout(guides = "collect", 
+              axes = 'collect')
+
+```
+
+```{r}
+block_section(
+  prop_section(
+    page_size = page_size(orient = "landscape"),
+    page_margins = page_mar(bottom = 1/8, top = 1/8,
+                            right = 1/4, left = 1/4),
+    type = "continuous"
+  )
+)
+```

### 2025-04-04 R/dpp_merge.R

@@ -25,6 +25,10 @@ dpp_merge <- function(dpp_components, daysrand_max = 0) {
 
  # include forms data (don't worry about filtering to visit/daysrand)
 
+ # blood pressure
+
+ # dpp_bp <- dpp_components$form$s03
+
  # education
  dpp_edu <- dpp_components$form$s06 %>%
   transmute(

### 2025-04-04 R/mesa_exclude.R

@@ -3,15 +3,21 @@
 
 mesa_exclude <- function(mesa_init) {
 
+
+  # drop_na(time_diabetes, status_diabetes) %>%
+
   # don't use nrow because the mesa data were pre-filtered
-  n_participants <- 6814
+  n_participants <- length(unique(mesa_init$id))
 
   e1 <- mesa_init %>%
-      # drop those without pre-diabetes
+    drop_na(time_diabetes, status_diabetes) %>%
+    # drop those without pre-diabetes
     filter(prediabetes == 1) %>%
     group_by(id) %>%
     slice(1) %>%
     ungroup()
+
+  stopifnot(all(e1$diabetes==0))
 
   n_participants <- c(n_participants, nrow(e1), NA_integer_)
 
@@ -22,10 +28,13 @@ mesa_exclude <- function(mesa_init) {
 
 
   out <- list(included = e1,
-              excluded = setdiff(mesa_init, e1))
+              excluded = mesa_init %>%
+                group_by(id) %>%
+                slice(1) %>%
+                filter(!id%in%e1$id))
 
   # assert no loss of participants
-  stopifnot(nrow(bind_rows(out)) == nrow(mesa_init))
+  stopifnot(n_participants[1] == nrow(out$excluded) + nrow(out$included))
 
   attr(out, 'exclusions') <- exclusions
 

### 2025-04-04 R/mesa_load.R

@@ -17,9 +17,10 @@ mesa_load <- function() {
   )
 
   mesa_visit_one <-
-    read_sas(file.path(fpath, "mesaex1_nodiabv20231128.sas7bdat")) %>%
+    read_sas(file.path(fpath, "mesaex1_diabv20240601.sas7bdat")) %>%
     transmute(
       id = idno,
+      diabetes = anydiabex1,
       prediabetes = prediabex1,
       visit = 1L,
       time_diabetes = exdiabpyr,
@@ -30,9 +31,9 @@ mesa_load <- function() {
       sex = factor(female, levels = c(0,1), labels = c("male", "female")),
       race_ethnicity = factor(race1c,
                               labels = c("caucasian",
+                                         "other",
                                          "african_american",
-                                         "hispanic",
-                                         "other")),
+                                         "hispanic")),
       education = factor(edugrp,
                          levels = c(0, 1, 2, 3),
                          labels = c("less than high school",
@@ -50,14 +51,14 @@ mesa_load <- function() {
       bmi = bmi1c
     )
 
+  # older version: mesaex2_nodiab.sas7bdat
+
   mesa_visit_two <-
-    read_sas(file.path(fpath, "mesaex2_nodiab.sas7bdat")) %>%
+    read_sas(file.path(fpath, "mesaex2_diabv20240601.sas7bdat")) %>%
     transmute(
       id = idno,
-      prediabetes = glucos2c >= 100 &
-        glucos2c <= 125 &
-        hba1c2 < 6.5    &
-        hba1c2 > 5.7,
+      diabetes = anydiabex2,
+      prediabetes = prediabex2,
       visit = 2L,
       time_diabetes = exdiabpyr26,
       status_diabetes = incdiab26,
@@ -83,7 +84,14 @@ mesa_load <- function() {
 
   bind_rows(mesa_visit_one,
             mesa_visit_two) %>%
-    drop_na(time_diabetes, status_diabetes) %>%
     arrange(id, visit)
 
 }
+
+# mb <- bind_rows(mesa_visit_one,
+#                 mesa_visit_two) %>%
+#   filter(prediabetes == TRUE) %>%
+#   drop_na(time_diabetes, status_diabetes)
+#
+# nrow(mb)
+

### 2025-04-04 R/parse_model_eqn_terms.R

@@ -1,0 +1,33 @@
+parse_model_eqn_terms <- function(data, labels){
+
+  term_type <- data$interaction[1]
+
+  res <- data %>%
+    mutate(
+      variable = recode(variable, !!!labels),
+      estimate = format(round(estimate, 8), nsmall = 8),
+    )
+
+  if(term_type == 'main'){
+    out <- res %>%
+      mutate(
+        eqn_term = case_when(
+          is.na(level) ~ glue("({variable}) * {estimate}"),
+          TRUE ~ glue("({variable} = {level}) * {estimate}")
+        )) %>%
+      pull(eqn_term) %>%
+      glue_collapse(sep = ' + ')
+  } else {
+    out <- res %>%
+      mutate(
+        eqn_term = case_when(
+          level == 'base' ~ glue("{variable} * [{estimate}"),
+          level == 'metformin' ~ glue(" + ({variable} = {level}) * {estimate}) + "),
+          TRUE ~ glue("({variable} = {level}) * {estimate})]")
+        )
+      ) %>%
+      pull(eqn_term) %>%
+      glue_collapse(sep = '')
+  }
+
+}

### 2025-04-04 R/tabulate_characteristics.R

@@ -7,6 +7,14 @@
 #' @param mesa_excluded
 tabulate_characteristics <- function(dpp_excluded, mesa_excluded, labels) {
 
+  # deal with the extra rows in mesa.
+
+  # mesa_excluded$excluded <- mesa_excluded$excluded %>%
+  #   arrange(id, visit) %>%
+  #   group_by(id) %>%
+  #   slice(1) %>%
+  #   filter(!(id %in% mesa_excluded$included$id))
+
   map2(
     .x = dpp_excluded,
     .y = mesa_excluded,
@@ -15,6 +23,10 @@ tabulate_characteristics <- function(dpp_excluded, mesa_excluded, labels) {
   )
 
 }
+
+# dpp_data <- tar_read(dpp_excluded)$excluded
+# mesa_data <- tar_read(mesa_excluded)$excluded
+# labels <- tar_read(labels)
 
 tabulate_characteristics_worker <- function(dpp_data, mesa_data, labels){
 
@@ -58,7 +70,7 @@ tabulate_characteristics_worker <- function(dpp_data, mesa_data, labels){
   }
 
   out <- tbl_summary(data = tbl_data,
-                     label = .label,
+                     label = as.list(.label),
                      by = 'study',
                      missing = 'no',
                      statistic = list(age_yrs = "{mean} ({sd})",

### 2025-04-04 R/tabulate_characteristics_optim.R

@@ -1,0 +1,129 @@
+#' .. content for \description{} (no empty lines) ..
+#'
+#' .. content for \details{} ..
+#'
+#' @title
+#' @param dpp_excluded
+#' @param mesa_excluded
+#' @param labels
+#' @param analysis
+#' @return
+#' @author bcjaeger
+#' @export
+tabulate_characteristics_optim <- function(data_analysis,
+                                           labels,
+                                           analysis) {
+
+
+  preds_met <- predict(analysis$fits$individualized,
+                       new_data = mutate(data_analysis,
+                                         treatment = "metformin"),
+                       type = 'survival',
+                       eval_time = 3) %>%
+    unnest(.pred)
+
+  preds_ili <- predict(analysis$fits$individualized,
+                       new_data = mutate(data_analysis,
+                                         treatment = "lifestyle"),
+                       type = 'survival',
+                       eval_time = 3) %>%
+    unnest(.pred)
+
+  inc_data <- data_analysis %>%
+    mutate(pred_risk_met = 1 - preds_met$.pred_survival,
+           pred_risk_ili = 1 - preds_ili$.pred_survival,
+           tx_optim = if_else(pred_risk_met < pred_risk_ili,
+                              "metformin", "lifestyle"))
+
+  incidence_tx <- cuminc(ftime = inc_data$time_diabetes,
+                         fstatus = inc_data$status_diabetes,
+                         group = inc_data$treatment)
+
+  inc_data_tx_optim <- filter(inc_data, tx_optim == treatment)
+
+  incidence_tx_optim <- cuminc(ftime = inc_data_tx_optim$time_diabetes,
+                               fstatus = inc_data_tx_optim$status_diabetes)
+
+  nnt_data <- incidence_tx[1:3] %>%
+    c(incidence_tx_optim[1]) %>%
+    map_dfr(~bind_cols(est=.x$est, time=.x$time), .id = 'group') %>%
+    group_by(group) %>%
+    mutate(tdiff = time - 3) %>%
+    filter(tdiff > 0) %>%
+    arrange(tdiff) %>%
+    slice(1) %>%
+    ungroup() %>%
+    select(group, est) %>%
+    mutate(group = str_remove(group, " 1$"),
+           group = recode(group, "1" = "model"),
+           arr = est[group == 'placebo'] - est,
+           nnt = 1 / arr) %>%
+    filter(group != 'placebo') %>%
[…]

### 2025-04-04 R/tabulate_cph.R

@@ -1,0 +1,158 @@
+
+tabulate_cph <- function(analysis, labels, pred_horizon = 3) {
+
+  rspec <- round_spec() %>%
+    round_using_decimal(digits = 2)
+
+  data_fit <- analysis$data_internal
+
+  data_recipe <- recipe(data_fit, time + status ~ .) %>%
+    step_rm(individualized, standard) %>%
+    update_role(race_ethnicity, new_role = 'id') %>%
+    step_impute_mean(all_numeric_predictors()) %>%
+    step_impute_mode(all_nominal_predictors()) %>%
+    step_scale(all_numeric_predictors())
+
+  data_prepped <- prep(data_recipe)
+
+  train <- juice(data_prepped)
+
+  cph <- coxph(
+    formula = Surv(time, status) ~ sex + hba1c_percent + chol_trig_mgdl +
+      treatment * (age_yrs + glucose_fasting_mgdl + bmi),
+    data = train,
+    model = TRUE,
+    x = TRUE
+  )
+
+  betas <- coef(cph)
+
+  lpm <- cph$means
+
+  bhaz <- suppressWarnings(basehaz(cph)) %>%
+    filter(time <= pred_horizon) %>%
+    slice(n()) %>%
+    pull(hazard)
+
+  continuous_pred_sds <- data_prepped$steps[[4]]$sds
+
+  lp_mean <- lpm %*% betas
+
+  # assert that the prediction equation we use will give the same answer
+  # as the predictRisk function from riskRegression
+  test_data <- bake(data_prepped, new_data = analysis$data_external)
+
+  pred_vars <- setdiff(all.vars(cph$formula), c("time", "status"))
+  pred_form <- as.formula( paste("~", as.character(terms(cph$formula))[3]) )
+  lp_data <-  select(test_data, all_of(pred_vars))
+
+  lp <-  model.matrix(pred_form, data = lp_data)[, -1L] %*% betas
+
+  preds_eqn <- 1 - exp(exp(lp-as.numeric(lp_mean)) * -bhaz)
+
+  preds_reg <- predictRisk(cph,
+                           newdata = test_data,
+                           times = pred_horizon)
+
+  # throws an error if any test predictions don't match prediction equation
+  stopifnot(
+    all(abs(as.numeric(preds_reg) - as.numeric(preds_eqn)) < 1e-10)
+  )
+
+  cph_eqn <- tidy(cph) %>%
+    separate(term, into = c("term", "interaction"),
[…]

### 2025-04-04 R/tabulate_cuminc.R

@@ -7,6 +7,7 @@
 tabulate_cuminc <- function(data_analysis, pred_horizon = 3) {
 
   data_cr <- data_analysis %>%
+    drop_na(time_death, status_death) %>%
     mutate(
       status = if_else(
         time_death < time_diabetes & status_death == 1,

### 2025-04-04 R/tabulate_optimal_tx.R

@@ -1,0 +1,60 @@
+
+tabulate_optimal_tx <- function(analysis, pred_horizon = 3) {
+
+  data_internal <- map(
+    .x = set_names(levels(analysis$data_internal$treatment)),
+    ~ mutate(analysis$data_internal, treatment = .x,
+             id_tmp = paste0('internal_', seq(n())))
+  )
+
+  data_external <- map(
+    .x = set_names(levels(analysis$data_external$treatment)),
+    ~ mutate(analysis$data_external, treatment = .x,
+             id_tmp = paste0('external_', seq(n())))
+  )
+
+  data <- map2(data_internal,
+               data_external,
+               ~bind_rows(internal = .x,
+                          external = .y,
+                          .id = 'type'))
+
+
+  results <- map(
+    .x = data,
+    ~ augment(analysis$fits$individualized,
+              new_data = .x,
+              type = 'survival',
+              eval_time = pred_horizon) %>%
+      unnest(cols = .pred) %>%
+      select(type, id_tmp, .pred_survival, treatment)
+  )
+
+  preds <- bind_rows(results) %>%
+    pivot_wider(names_from = treatment, values_from = .pred_survival) %>%
+    rowwise() %>%
+    mutate(tx_opt = which.max(c(lifestyle, metformin))) %>%
+    ungroup() %>%
+    mutate(tx_opt = factor(tx_opt,
+                           levels = c(1, 2, 3),
+                           labels = c('lifestyle', 'metformin', 'placebo')))
+
+  optim_risk <- preds %>%
+    pivot_longer(cols = c(lifestyle, metformin, placebo), names_to = 'tx_cf') %>%
+    group_by(type, tx_opt, tx_cf) %>%
+    summarize(risk_mean = mean(100 * (1-value)),
+              risk_sd = sd(100 * (1-value))) %>%
+    mutate(tbv = table_glue("{risk_mean} ({risk_sd})")) %>%
+    select(-starts_with("risk")) %>%
+    pivot_wider(names_from = tx_cf, values_from = tbv)
+
+  optim_counts <- preds %>%
+    group_by(type) %>%
+    count(tx_opt) %>%
+    mutate(p = n / sum(n)) %>%
+    transmute(type, tx_opt, tbn = table_glue("{n} ({100 * p}%)"))
+
+  left_join(optim_counts, optim_risk)
+
+
+}

### 2025-04-04 R/tabulate_reclassification.R

@@ -1,0 +1,52 @@
+
+
+
+tabulate_reclassification <- function(analysis) {
+
+
+  risk_cuts = c(0, 0.1, 0.2, 1)
+
+  data <- bind_rows(internal = analysis$data_internal,
+                    external = analysis$data_external,
+                    .id = 'set') %>%
+    mutate(
+      across(
+        .cols = c(individualized, standard),
+        .fns = ~ cut(.x,
+                     breaks = risk_cuts,
+                     include.lowest = TRUE,
+                     right = FALSE,
+                     labels = c("0 to < 10%",
+                                "10% to < 20%",
+                                "\u2265 20%"))
+      ),
+      case = if_else(time < 3 & status == 1, 'yes', 'no'),
+      set = factor(set, levels = c("internal", "external"))
+    )
+
+  by_set <- data %>%
+    group_by(set) %>%
+    count(individualized, standard) %>%
+    mutate(n = table_glue("{n} ({100 * n / sum(n)}%)")) %>%
+    pivot_wider(names_from = individualized, values_from = n,
+                values_fill = "0")
+
+  by_sex <- data %>%
+    group_by(sex) %>%
+    count(individualized, standard) %>%
+    mutate(n = table_glue("{n} ({100 * n / sum(n)}%)")) %>%
+    pivot_wider(names_from = individualized, values_from = n,
+                values_fill = "0")
+
+  by_race <- data %>%
+    group_by(race_ethnicity) %>%
+    count(individualized, standard) %>%
+    mutate(n = table_glue("{n} ({100 * n / sum(n)}%)")) %>%
+    pivot_wider(names_from = individualized, values_from = n,
+                values_fill = "0")
+
+  list(set = by_set,
+       sex = by_sex,
+       race = by_race)
+
+}

### 2025-04-04 R/validate_preds.R

@@ -6,7 +6,7 @@
 #' @param data_analysis
 validate_preds <- function(data_analysis,
                            pred_horizon = 3,
-                           risk_cut = 0.125) {
+                           risk_cut = 0.2) {
 
   initial_split <- data_analysis %>%
     mutate(surv = Surv(time_diabetes, status_diabetes == 1)) %>%
@@ -31,9 +31,14 @@ validate_preds <- function(data_analysis,
   derivation_folds <- vfold_cv(derivation)
 
   rec_standard <- recipe(surv ~ ., data = derivation) %>%
-    step_impute_mean(all_numeric_predictors()) %>%
-    step_impute_mode(all_nominal_predictors()) %>%
-    update_role(race_ethnicity, new_role = 'id')
+    update_role(race_ethnicity, new_role = 'id') %>%
+    step_impute_knn(all_predictors())
+
+  rec_reviewer_1 <- derivation %>%
+    select(surv, glucose_fasting_mgdl, hba1c_percent, race_ethnicity) %>%
+    recipe(surv ~ ., data = .) %>%
+    update_role(race_ethnicity, new_role = 'id') %>%
+    step_impute_knn(all_predictors())
 
   rec_interaction <- rec_standard %>%
     step_interact(terms = ~treatment:matches("age_yrs|bmi|^glucose"))
@@ -46,12 +51,17 @@ validate_preds <- function(data_analysis,
     add_recipe(rec_standard) %>%
     add_model(cph_spec)
 
+  wf_reviewer_1 <- workflow() %>%
+    add_recipe(rec_reviewer_1) %>%
+    add_model(cph_spec)
+
   wf_individualized <- workflow() %>%
     add_recipe(rec_interaction) %>%
     add_model(cph_spec)
 
   wf <- list(
     standard = wf_standard,
+    reviewer_1 = wf_reviewer_1,
     individualized = wf_individualized
   )
 
@@ -76,10 +86,10 @@ validate_preds <- function(data_analysis,
       group_by(model) %>%
       mutate(.id = seq(n()))
 
-    test_out[[i]] <- tibble(time = test$surv[,1],
-                            status = test$surv[,2],
-                            sex = test$sex,
-                            race_ethnicity = test$race_ethnicity)
+    test_out[[i]] <- test %>%
+      mutate(time = surv[,1],
+             status = surv[,2]) %>%
+      select(-surv)
 
   }
 
[…]

### 2025-04-04 R/visualize_calib.R

@@ -1,0 +1,99 @@
+
+visualize_calib <- function(analysis,
+                            pred_horizon = 3,
+                            type = 'external') {
+
+  data_calib <- switch(type,
+                       'internal' = analysis$data_internal,
+                       'external' = analysis$data_external)
+
+  groups <- list(
+    overall = data_calib,
+    men = filter(data_calib, sex == 'male'),
+    women = filter(data_calib, sex == 'female'),
+    black = filter(data_calib, race_ethnicity == 'african_american'),
+    white = filter(data_calib, race_ethnicity == 'caucasian'),
+    hispanic = filter(data_calib, race_ethnicity == "hispanic"),
+    other = filter(data_calib, race_ethnicity == "other")
+  )
+
+  output <- map(
+    groups,
+
+    .f = ~ {
+
+      .scalib <-
+        scalib(pred_risk = .x$individualized,
+               pred_horizon = pred_horizon,
+               event_status = .x$status,
+               event_time = .x$time)
+
+
+      # keeping max dimension low to reduce computation time
+      .scalib_slope <- scalib_hare(scalib_object = .scalib,
+                                   hare_linear_time = TRUE,
+                                   hare_linear_risk = TRUE)
+
+
+      # use 7 risk groups. default is 10, but 7 looks nicer on the final plot
+
+      .scalib_test <- scalib_gnd(scalib_object = .scalib,
+                                 group_count_init = 10)
+
+      data_pointrange <- .scalib_test %>%
+        getElement('data_outputs') %>%
+        unnest(gnd_data)
+
+
+      gg_data <- .scalib_slope %>%
+        getElement("data_outputs") %>%
+        select(._id_., hare_data_plot) %>%
+        unnest(hare_data_plot)
+
+      fig_cal_slope <- ggplot(gg_data) +
+        aes(x = predicted, y = observed) +
+        geom_pointrange(
+          data = data_pointrange,
+          mapping = aes(x = percent_expected,
+                        y = percent_observed,
+                        ymin = percent_observed - 1.96 * sqrt(variance),
+                        ymax = percent_observed + 1.96 * sqrt(variance)),
+          color = 'grey') +
+        geom_line() +
+        geom_abline(col = 'grey', linetype = 2, intercept = 0, slope = 1) +
+        scale_x_continuous(limits = c(0,1),
+                           # expand = c(0,0),
+                           breaks = seq(0, 1, by = 0.2),
[…]

### 2025-04-04 R/visualize_dcurve.R

@@ -1,0 +1,84 @@
+
+visualize_dcurve <- function(analysis, pred_horizon = 3) {
+
+  library(dcurves)
+
+  dca_objects <- map(
+    .x = analysis[c('data_internal', 'data_external')],
+    .f = ~ {
+
+      init <- dca(formula = Surv(time, status) ~ individualized + standard,
+                  data = .x,
+                  time = 3,
+                  thresholds = seq(0.01, 0.50, length.out = 100))
+
+      gg_data <- init$dca %>%
+        mutate(label = str_to_title(label))
+
+      net_ppl <- gg_data %>%
+        filter(variable != 'all') %>%
+        filter(threshold > 0.19,
+               threshold < 0.21) %>%
+        group_by(label) %>%
+        summarize(net_ppl = round(100 * mean(net_benefit), digits = 0)) %>%
+        filter(label == "Individualized") %>%
+        deframe()
+
+      sl_label <- glue(
+        "At 20%, the individualized model
+        identifies {net_ppl} patients who will
+        develop T2 diabetes in the next 3
+        years per 100 in target population."
+      )
+
+      arrow_end <- gg_data %>%
+        filter(threshold > 0.19,
+               threshold < 0.21,
+               variable == "individualized") %>%
+        group_by(variable) %>%
+        summarize(threshold = mean(threshold),
+                  net_benefit = mean(net_benefit))
+
+      fig <- ggplot(gg_data) +
+        aes(x = threshold,
+            y = net_benefit,
+            color = label) +
+        geom_line(linewidth = 3/4,
+                  show.legend = c(color = TRUE, alpha = FALSE)) +
+        coord_cartesian(ylim = c(0, 0.125)) +
+        theme_bw() +
+        theme(panel.grid = element_blank(),
+              legend.position = 'top') +
+        labs(x = "Threshold, %",
+             y = "Net benefit",
+             color = "") +
+        scale_x_continuous(labels = scales::percent) +
+        scale_color_manual(
+          values = c("deeppink",
+                     "seagreen",
+                     "black",
+                     "grey80")
+        ) +
+        annotate('text',
+                 label = sl_label,
+                 x = 0.375,
+                 y = 0.095) +
+        annotate(
+          'curve',
+          x = .25, # Play around with the coordinates until you're satisfied
+          y = 0.08,
+          yend = arrow_end$net_benefit + 0.001,
[…]

### 2025-04-04 _targets.R

@@ -1,4 +1,10 @@
 
+# TODO: redo table 1 using predicted best tx as stratifier in dpp
+# TODO: re-read and clarify methods section
+# TODO: NNT to prevent 1 incident DM using ILI only, MF only, and
+#       using recommendation from model
+# TODO: add 1 sentence description of cstat calib and nri
+
 source("packages.R")
 source("conflicts.R")
 
@@ -10,7 +16,7 @@ library(future.callr)
 plan(callr)
 
 
-manuscript_version <- 1
+manuscript_version <- 7
 
 if(!dir.exists(glue("doc/manuscript-v{manuscript_version}"))){
   dir.create(glue("doc/manuscript-v{manuscript_version}"))
@@ -20,6 +26,7 @@ if(!dir.exists(glue("doc/manuscript-v{manuscript_version}"))){
 labels_tar <- tar_target(
   labels,
   list(
+
     variables = c(
       age_yrs = "Age",
       race_ethnicity = "Race/ethnicity",
@@ -33,7 +40,11 @@ labels_tar <- tar_target(
       chol_trig_mgdl = "Triglycerides",
       chol_ldlc_mgdl = "Low-density lipoprotein cholesterol",
       chol_hdl_mgdl = "High-density lipoprotein cholesterol",
-      bmi = "Body mass index"),
+      bmi = "Body mass index",
+      sbp = "Systolic blood pressure",
+      dbp = "Diastolic blood pressure"
+    ),
+
     units = c(
       age_yrs = "years",
       glucose_fasting_mgdl = "mg/dl",
@@ -41,19 +52,24 @@ labels_tar <- tar_target(
       chol_trig_mgdl = "mg/dl",
       bmi = "kg/m2",
       chol_ldlc_mgdl = "mg/dl",
-      chol_hdl_mgdl = "mg/dl"
+      chol_hdl_mgdl = "mg/dl",
+      sbp = "mm Hg",
+      dbp = "mm Hg"
     ),
+
     levels = list(
       sex = c(
         "male" = "Male",
         "female" = "Female"
       ),
+
       race_ethnicity = c(
         "caucasian" = "Non-Hispanic White",
         "african_american" = "Non-Hispanic Black",
         "hispanic" = "Hispanic",
         "other" = "Other/Chinese"
       ),
+
       education = c(
         "less than high school" = "< High School",
         "high school" = "High School Graduate",
[…]

### 2025-04-04 doc/review/reviewer_response_demo_cstat.R

@@ -1,0 +1,19 @@
+
+
+set.seed(123)
+
+n <- 10000
+
+data_common <- tibble(time = round(runif(n) * 100),
+                      x = runif(n, min = 0.5, max = 1),
+                      status = rbinom(n, size = 1, prob = x))
+
+data_rare <- tibble(time = round(runif(n) * 100),
+                      x = runif(n, min = 0, max = 0.5),
+                      status = rbinom(n, size = 1, prob = x))
+
+# outcome is more common => lower C of 0.562
+summary(coxph(Surv(time, status) ~ x, data = data_common))$concordance
+
+# outcome is more rare => Higher C of 0.656
+summary(coxph(Surv(time, status) ~ x, data = data_rare))$concordance

### 2025-04-04 packages.R

@@ -26,6 +26,7 @@ library(table.glue)
 library(data.table)
 library(flextable)
 library(cmprsk)
+library(survIDINRI)
 
 
 # standardizes starting location for filepaths
